---
title: "Live Session 11"
author: "Tracie Scott"
date: "November 28, 2016"
output: 
  html_document:
    keep_md: true
---
# Tracie Scott - Live Session 11 
# Using hSales data

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
# Display session info
sessionInfo()
# load any packages required
library(fpp) #fpp package must be installed first

# SET WORKING DIRECTORY HERE!!
setwd("C:/Users/User/Documents/hsales")

# remember that the working directory gets reset between chunks
```

# Problem: Time Series
####Plot the time series. 
####Can you identify seasonal fluctuations and/or a trend? 
## Plot for Time Series

```{r A-TimeSeries, echo=TRUE}
library(fpp) #fpp package must be installed first
hsales2 <- ts(hsales,start=c(1973,1),frequency=12)
plot(hsales2)
```

## Response for Time Series
#### Notice repeated spikes in the data of similar magnitude that could be seasonal fluctuations. 
#### While the data is fluctuating over time, there is not a consistent trend.
#### .  
# Problem : Traditional Decommposition
#### Use a classical decomposition to calculate the trend-cycle and seasonal indices. 
#### Do the results support the graphical interpretation from part (a)?
## Plot for Traditional Decomposition

```{r B-Decompose, echo=TRUE}
fitd <- decompose(hsales)
plot(fitd)
```

## Response for Traditioanl Decomposition
#### Yes, this also seems to show a repeated seasonal component without a constant trend from start to finish. 
#### The decompose also shows a significant ‘random’ effect after the trend and seasonal components are removed.
#### .
# Problem:  Seasonally Adjusted
#### Compute and plot the seasonally adjusted data.
## Plot for Seasonally Adjusted

```{r C-Seasonal, echo=TRUE}
hsalesadj <- seasadj(fitd)
plot(hsalesadj)
```

## Response for Seasonally Adjusted
#### If this chart 'removes' the seasonal findings, then it appears that most of the fluctuations are due to 
#### random variations in the data as the seasonally adjusted chart looks much like the original data.
#### . 
# Problem: Create Outlier
#### Change one observation to be an outlier (e.g., add 500 to one observation)
#### Recompute the seasonally adjusted data. 
#### What is the effect of the outlier?
#### Does it make any difference if the outlier is near the end rather than in the middle of the time series?
## Plost for Outlier (mid-series)

```{r D-Outlier, echo=TRUE}
hsales3 <- ts(c(hsales[1:54],hsales[55]+200,hsales[56:275]),start=c(1973,1),frequency=12)
plot(hsales3)

fitd <- decompose(hsales3)
plot(fitd)

hsalesadj <- seasadj(fitd)
plot(hsalesadj)
```

## Response for Create Outlier
#### The outlier makes the seasonally adjusted chart much harder to interpret,
#### but it does not appear to have a significant impact on the analysis.
#### .
# Problem Decompose with STL
#### Now use STL to decompose the original series.
#### Do the results support the graphical interpretation from part (a)?
## Plot for STL

```{r E-stl, echo=TRUE}
fit <- stl(hsales, s.window=5)
plot(fit)
```

## Response for STL
#### The output from stl reflects more variability in the seasonal component than did decompose, 
#### but the results are similar. Both decomposition methods seem to indicate there is a repeated seasonal
#### pattern without a pronounced upward or downward trend over time. 
#### However, the ‘random’ and ‘residual’ plots show a great deal of unexplained variability in the data even after
#### removing the seasonal the trend components.
